---
title: "Statistical analysis for Cu quotas: eukaryotic algae versus het. bacteria"
author: "Ania"
date: "January 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Testing the differences in macronutrient ratios and metal quotas of eukaryotic algae and het. bacteria

Question: "Are metal quotas (Me:C) of bacteria and phytoplankton different?"

Null hypothesis : "Metal quotas (Me:C medians) of bacteria and phytoplankton do not differ"

Considerations:

* assess if data is normally distributed (assessed wity qqPlot in package car)
* assess if variances are equal Levene test for normally distributed data and F-test for non-normally distributed data
* bacteria datasets contain observations of 5 quotas for each species, therefore are not independent observations, pseudoreplicates *

### Statistical tests

Tests for homogeneity of variances:  

* Levene's test (for non-normally distributed data)
* F test (for normally distributed data)
* Bartlette test (sensitive to departures from normality)

Testing differences between two samples:  

* Wilcoxon Rank Sum test, also known as Mann-Whitney test. It is a nonparametric test of the null hypothesis that it is equally likely that a randomly selected value from one sample will be less than or greater than a randomly selected value from a second sample. Does not assume normal distribution

* Moods median test: two sample version of the Sign Test, is used to determine whether the median of two independent samples are equal. Since p-value = .0405 < .05 = α, we reject the null-hypothesis, and conclude there is a significant difference between the two population medians.

* Welch t-test, for unequal variances but assumes normal distribution

> Running the tests

## 1. Preparing the data

```{r message=FALSE}

# load the data from csv
library(tidyverse)
data <-read_csv("Data/06_Cu-quotas-lit-aquil-tidydata.csv") 

glimpse(data)

# sort out the variable formats
data$Strategy <- as.factor(data$Strategy)
data$Taxon<- as.factor(data$Taxon)
data$Species<- as.factor(data$Species)
data$Domain<- as.factor(data$Domain)
```


## 2 Subsetting
Note, there is a single, abnormal observation in the phytoplakton data @ Cu:C of 30, this is not included in the test.

Note about data from Sunda & Huntsman 1995: where Cu:C for E. huxleyi and T.oceanica were measured in triplicate, that data was averaged.

```{r message=FALSE}
data_no_cyano <- data%>% filter(Domain %in% c("Euk_algae","Het_bacteria"), Cu_C < 20)

```


## 3. Assessing normalitey. 

Note, data points fall outside the 95% confidence intervals suggesting non-normal distribution of data

```{r message=FALSE}

library(car)
qqPlot(data_no_cyano$Cu_C)

```

## 4. Testing for equality of variances

Variances are NOT equal as p = 0.019 for Levene Test, similarily with F test we get significant p>0.001

```{r message=FALSE}

leveneTest (Cu_C ~ Domain, data = data_no_cyano)

# or F-test if data were to be normally distributed
var.test(Cu_C~Domain, data=data_no_cyano)# - this is F-test
```

### 5. Non parametric tests

#####Wilcox Rank sum test (Mann-Whitney U test) and Moods median test (for further exploration). Result: since p-value for both Wilcox and Moods median is >0.05, we accept the null hypothesis that Cu quotas of bacteria and phytoplankton on this dataset are the same. Note this Wilcox test function in R is by default two.sided.

```{r message=FALSE}

wilcox.test(Cu_C~Domain, data = data_no_cyano)

library(RVAideMemoire) # contains function for the moods median test
mood.medtest(Cu_C ~ Domain,data = data_no_cyano)

```



